Re-examining cross-cultural similarity judgments using language statistics

### 2023-02-02 code/4_paper/manuscript/manuscript.Rmd
small fixes

[^1]: With the preregistered exclusion criterion, all three corpora were significant predictors.

@@

```{=tex}
\begin{CJK*}{UTF8}{gbsn}
Each participant completed all 105 triads in blocks of 21 trials at a time (10 test triads, 10 filler triads, and 1 attention check per page), by selecting the match most related to the cue. We elicited similarity match responses by asking "Which thing is most closely related to the bolded [cue] item?" In Mandarin Chinese, we used the phrasing for "most closely related" reported by Ji et al. (2004): "哪个事物与加粗项关联最紧密?" And in Vietnamese, we translated this as "Thứ nào liên quan nhất với thứ được in đậm?".
The test triads were presented with 105 filler triads mixed in, to obscure the taxonomic-thematic two-answer forced choice structure of the test stimuli and reduce the likelihood that participants would become aware of the design.
The filler triads were groups of three semantically related words, but where the match options were not distinguished by thematic vs. taxonomic similarity, for example, the cue "bird" with match options "lizard" and "toad." Additionally, we included 11 attention check trials, which were formatted similarly to the test and filler triads but included an instruction instead of a cue item, e.g., "Choose wife" with match options "wife" and "husband." In total, each participant completed 210 similarity judgments (105 test triads and 105 fillers) and 11 attention check questions, with all items presented in randomized orders that varied between subjects.
\end{CJK*}
```
## Corpus model

@@

[^5]: With the preregistered exclusion criteria, the English (EN) and Mandarin (ZH) corpora were significant predictors.

For Vietnamese participants' responses, only the Vietnamese (VI) and Mandarin (ZH) corpora were significant predictors[^6].
EN corpus: $\beta$ = `r round(fixef(fit.ENVIZH_VN_cos)["theme_cosine_prop_E"], 2)`, $\chi^2$(`r fit.ENVIZH_VN_cos.anova["theme_cosine_prop_E", "Df"]`) = `r round(fit.ENVIZH_VN_cos.anova["theme_cosine_prop_E", "Chisq"], 2)`, p `r ifelse( fit.ENVIZH_VN_cos.anova["theme_cosine_prop_E", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.ENVIZH_VN_cos.anova["theme_cosine_prop_E", "Pr(>Chisq)"], 3)), "< .001")`. VI corpus: $\beta$ = `r round(fixef(fit.ENVIZH_VN_cos)["theme_cosine_prop_V"], 2)`, $\chi^2$(`r fit.ENVIZH_VN_cos.anova["theme_cosine_prop_V", "Df"]`) = `r round(fit.ENVIZH_VN_cos.anova["theme_cosine_prop_V", "Chisq"], 2)`, p `r ifelse( fit.ENVIZH_VN_cos.anova["theme_cosine_prop_V", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.ENVIZH_VN_cos.anova["theme_cosine_prop_V", "Pr(>Chisq)"], 3)), "< .001")`. ZH corpus: $\beta$ = `r round(fixef(fit.ENVIZH_VN_cos)["theme_cosine_prop_M"], 2)`, $\chi^2$(`r fit.ENVIZH_VN_cos.anova["theme_cosine_prop_M", "Df"]`) = `r round(fit.ENVIZH_VN_cos.anova["theme_cosine_prop_M", "Chisq"], 2)`, p `r ifelse( fit.ENVIZH_VN_cos.anova["theme_cosine_prop_M", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.ENVIZH_VN_cos.anova["theme_cosine_prop_M", "Pr(>Chisq)"], 3)), "< .001")`.

[^6]: With the preregistered exclusion criteria, all three corpora were significant predictors.

>>>>

[^1]: We also carried out our analysis using raw lexical cooccurrences
and obtained similar results.

@@

```{=tex}
\begin{CJK*}{UTF8}{gbsn}
Each participant completed all 105 triads in blocks of 21 trials at a time (10 test triads, 10 filler triads, and 1 attention check per page), by selecting the match most related to the cue. We elicited similarity match responses by asking "Which thing is most closely related to the bolded [cue] item?" In Mandarin Chinese, we used the phrasing for "most closely related" reported by Ji et al. (2004): "关联". And in Vietnamese, we translated this as "liên quan nhất".
The test triads were presented with 105 filler triads mixed in, to obscure the taxonomic-thematic two-answer forced choice structure of the test stimuli and reduce the likelihood that participants would become aware of the design.
The filler triads were groups of three semantically related words, but where the match options were not distinguished by thematic vs. taxonomic similarity, for example, the cue "bird" with match options "lizard" and "toad." Additionally, we included 11 attention check trials, which were formatted similarly to the test and filler triads but included an instruction instead of a cue item, e.g., "Choose wife" with match options "wife" and "husband." In total, each participant completed 210 similarity judgments (105 test triads and 105 fillers) and 11 attention check questions, with all items presented in randomized orders that varied between subjects.
\end{CJK*}
```
## Corpus model

@@

[^2]: With the preregistered exclusion criterion, only the English
(EN) and Mandarin (ZH) corpus were significant predictors.

For Vietnamese participants' responses, only the Vietnamese (VI) and Mandarin (ZH) corpora were significant predictors[^3].
EN corpus: $\beta$ = `r round(fixef(fit.ENVIZH_VN_cos)["theme_cosine_prop_E"], 2)`, $\chi^2$(`r fit.ENVIZH_VN_cos.anova["theme_cosine_prop_E", "Df"]`) = `r round(fit.ENVIZH_VN_cos.anova["theme_cosine_prop_E", "Chisq"], 2)`, p `r ifelse( fit.ENVIZH_VN_cos.anova["theme_cosine_prop_E", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.ENVIZH_VN_cos.anova["theme_cosine_prop_E", "Pr(>Chisq)"], 3)), "< .001")`. VI corpus: $\beta$ = `r round(fixef(fit.ENVIZH_VN_cos)["theme_cosine_prop_V"], 2)`, $\chi^2$(`r fit.ENVIZH_VN_cos.anova["theme_cosine_prop_V", "Df"]`) = `r round(fit.ENVIZH_VN_cos.anova["theme_cosine_prop_V", "Chisq"], 2)`, p `r ifelse( fit.ENVIZH_VN_cos.anova["theme_cosine_prop_V", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.ENVIZH_VN_cos.anova["theme_cosine_prop_V", "Pr(>Chisq)"], 3)), "< .001")`. ZH corpus: $\beta$ = `r round(fixef(fit.ENVIZH_VN_cos)["theme_cosine_prop_M"], 2)`, $\chi^2$(`r fit.ENVIZH_VN_cos.anova["theme_cosine_prop_M", "Df"]`) = `r round(fit.ENVIZH_VN_cos.anova["theme_cosine_prop_M", "Chisq"], 2)`, p `r ifelse( fit.ENVIZH_VN_cos.anova["theme_cosine_prop_M", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.ENVIZH_VN_cos.anova["theme_cosine_prop_M", "Pr(>Chisq)"], 3)), "< .001")`.

[^3]: With the preregistered exclusion criteria, all three corpora were significant predictors.

### 2023-02-02 code/4_paper/manuscript/manuscript.tex
small fixes

Data collection and analyses for this study were pre-registered, and the
pre-registration is available at {[}blinded for peer review{]}. We
recruited 200 participants from the US, 200 from mainland China, and 199
from Vietnam. All participants were recruited through snowball sampling,
which, in the US was seeded with student email lists at a large
university, in China (CN) with student groups on WeChat, and in Vietnam
(VN) with Vietnam-based student groups on Facebook. US participants were
compensated with \$5 gift certificates (USD), CN participants received
25CNY through WeChat credit transfer, and VN participants received
50,000₫ (VND) in phone credit,

@@

All stimuli and scripts for the experiment and analysis are available at
{[}BLINDED FOR REVIEW{]}. We collected and adapted stimuli from previous
studies to create test triads consisting of a cue with one thematic and
one taxonomic match option. For example, ``cow,'' ``grass,'' and
``chicken,'' where ``cow'' is the cue, ``grass'' is the thematic match,
and ``chicken'' the taxonomic match. We included 105 such triads, a
superset including triads pulled from supplemental information and
in-text examples across the literature, and others that we adapted or
created . We selected triads on the basis of cultural familiarity within
the US, Vietnam, and China. The triads were translated from English to
Vietnamese and Mandarin by fluent bilingual speakers of each language,
and back-translated to English by another fluent bilingual who was naive
to the original English version. The back-translations were checked
against the original English to identify ambiguities in the translation,
which were resolved through discussion and selection of alternative
terms when relevant.

@@

\begin{CJK*}{UTF8}{gbsn}
Each participant completed all 105 triads in blocks of 21 trials at a time (10 test triads, 10 filler triads, and 1 attention check per page), by selecting the match most related to the cue (following Ji et al., 2004): "Which thing is most closely related to the bolded item?" / "哪个事物与加粗项关联最紧密?" / "Thứ nào liên quan nhất với thứ được in đậm?".
The test triads were presented with 105 filler triads mixed in, to obscure the taxonomic-thematic two-answer forced choice structure of the test stimuli and reduce the likelihood that participants would become aware of the design.
The filler triads were groups of three semantically related words, but where the match options were not distinguished by thematic vs. taxonomic similarity, for example, the cue "bird" with match options "lizard" and "toad." Additionally, we included 11 attention check trials, which were formatted similarly to the test and filler triads but included an instruction instead of a cue item, e.g., "Choose wife" with match options "wife" and "husband." In total, each participant completed 210 similarity judgments (105 test triads and 105 fillers) and 11 attention check questions, with all items presented in randomized orders that varied between subjects.
\end{CJK*}

@@

In sum, we replicated the differences documented by Ji et al. (2004)
between the US and China. We did not find that Vietnamese participants
grouped particularly with Chinese participants, however. Instead they
were intermediate between the two. Accordingly, we did not find strong
support for the hypothesis that there are overall biases toward thematic
responding across Asian cultural contexts broadly.

@@

First, to test whether variation in language statistics can explain
differences in similarity judgments, we use a mixed-effects logistic
regressions fit individually for each country. Our models predicted
responses by individual participants to particular triads (0=taxonomic
or 1=thematic) with fastText predictions (proportion of cosine distance)
as a fixed effect and participant and triad as random effects. If
language statistics, contribute to the differences in similarity
judgments, we would expect each language corpus to be a good predictor
for similarity judgment responding in its corresponding context.

@@

If language statistics are able to predict meaningful culture-specific
variation in similarity judgments (rather than just consistency across
cultures), we would expect each corpus to be the best predictor of its
corresponding culture compared to the other two corpora. We directly
compared the corpus models by including all three corpus predictors as
fixed effects in three mixed-effect regressions (predicting US, Vietnam
and China responding) with the same random effects as above.

For US responding: only the English (EN) corpus was a significant
predictor\footnote{With the preregistered exclusion criterion, only the
  English (EN) and Mandarin (ZH) corpus were significant predictors.}.
EN corpus: \(\beta\) = -6.96, \(\chi^2\)(1) = 16.57, p \textless{} .001.
VI corpus: \(\beta\) = -2.22, \(\chi^2\)(1) = 3.73, p = 0.054. ZH
corpus: \(\beta\) = -3.18, \(\chi^2\)(1) = 3.37, p = 0.066.

For Vietnam responding, only the Vietnamese (VI) and Mandarin (ZH)
corpus were significant predictors\footnote{With the preregistered
  exclusion criterion, all three corpora were significant predictors.}.
EN corpus: \(\beta\) = -2.98, \(\chi^2\)(1) = 2.58, p = 0.108. VI
corpus: \(\beta\) = -2.75, \(\chi^2\)(1) = 4.84, p = 0.028. ZH corpus:
\(\beta\) = -4.39, \(\chi^2\)(1) = 5.49, p = 0.019.

For China responding, only the Mandarin (ZH) and English (EN) corpus
were significant predictors. EN corpus: \(\beta\) = -3.32, \(\chi^2\)(1)
= 7.18, p = 0.007. VI corpus: \(\beta\) = -1.59, \(\chi^2\)(1) = 3.63, p
= 0.057. ZH corpus: \(\beta\) = -5.31, \(\chi^2\)(1) = 17.69, p
\textless{} .001.

We observed some language specificity in this analysis (Figure 2). The
English corpus was the best predictor for US responding, and the
Mandarin corpus was the best predictor for Chinese responding. While
this is not the case with the Vietnamese corpus and the Vietnam
responding, the Vietnamese corpus was still a significant predictor for
the Vietnam responding. These results support our hypothesis that
specific language statistics can predict cross-cultural variations of
similarity judgment.

@@

However, in all cultural contexts, adding the other two corpora produced
a significantly better fit than the identical model with only the
corresponding corpus included as a predictor (US response: \(\chi^2\)(2)
= 7.93, p = 0.019; Vietnam response: \(\chi^2\)(2) = 13.72, p = 0.001;
China response: \(\chi^2\)(2) = 10.56, p = 0.005). This analysis
suggests that culture-specific inputs to similarity judgment (as proxied
by language statistics) do not fully explain cross-cultural differences
in similarity judgment, whether due to noise in the estimates of
language statistics or true additional variance picked up by embeddings
from other corpora.

@@

One possible concern with our approach is that our models are only able
to account for variation in similarity judgment because the word
embedding models pick up on cross-cultural differences in the
conceptualization of similarity. Perhaps some aspect of the geometry of
the embedding space specifically reflects these differences rather than
capturing more item-specific patterns of co-occurrence.

If this is the case, the embedding models should be less accurate at
predicting responding for filler items (such as whether ``tomorrow'' or
``yesterday'' is more related to ``today'') because these items do not
have a thematic/taxonomic structure. Additionally, even when one option
might be more related to the cue, that relationship is not systematic
throughout the set of fillers. On the other hand, if the embedding
models capture similarity judgments more generally (beyond
thematic/taxonomic preferences), they should predict judgments for the
unstructured filler items as well as they do for our experimental items.

For each filler triad, we randomly assigned one of the responding
options as `Word1'. Running a mixed-effects logistic regression to
predict responding (Word1 or Word2) with country as a fixed effect and a
random effect structure as above Q1, we found no effect of country on
filler responding (\(\chi^2\)(2) = 0.74, p = 0.692).

Using the same mixed-effects logistic regression structure as the single
corpus models in Q2, we predicted responses (1=word1 or 0=word2) in each
cultural context with the corresponding corpus predictor as a fixed
effect, and participant and triad as random effects. In all cultural
contexts, the corresponding corpus was a significant predictor of
responding, with p \textless{} 0.05 and \(\beta\) from -9.59 to -3.14.
These results show that the word embedding models predict general
similarity in addition to structured thematic/taxonomic similarity.

To investigate whether language statistics predict taxonomic/thematic
triads and filler items differently, we compared the variance accounted
for by the models of the two different types of items. We ran two
mixed-effects logistic regression models that predict responding across
cultural contexts (1=thematic or word1, 0=taxonomic or word2), with the
corresponding corpus predictor as a fixed effect and participant and
item as random effects. Consistent with our results above, corresponding
corpus was a significant predictor for responding to both triads and
filler items (triads: \(\beta\) = -1.79, \(\chi^2\)(1) = 75.58, p
\textless{} .001, fillers: \(\beta\) = -2, \(\chi^2\)(1) = 70.89, p
\textless{} .001). Importantly, we found comparable conditional \(R^2\)
values when the corresponding corpus was used to predict only triad
items (\(R^2\) = 0.3) and only filler items (\(R^2\) = 0.26). This
result provides evidence for our view that word embeddings index general
tendencies in similarity judgment, beyond a specific dimension of
thematic/taxonomic relations.

@@

Do members of different cultures vary in their conceptualization of
similarity? Cross-cultural differences in similarity judgment would seem
to suggest this conclusion. In this paper, we consider whether
statistics of the environment (as indexed by language statistics) can
account for cross-cultural differences in a classic similarity judgment
paradigm. We replicated the previously documented contrast between
English speakers in the US and Mandarin Chinese speakers from East Asia
(mainland China, Taiwan, Hong Kong, and Singapore): Chinese participants
preferred thematic relations to a greater extent compared to US
participants. Our sample of Vietnamese participants showed an
intermediate preference but were not significantly different from either
Chinese or US participants. This finding suggests some limitations on
the generality of the cultural account, which proposes that
thematic/taxonomic similarity preference aligns with East Asian/Western
tendency for holistic vs analytic processing.

In contrast, we found some support for the environmental statistics
account: word embeddings based on language specific corpora were a good
predictor for the corresponding country's similarity judgments, even
when other corpus statistics were included, and even with filler triads
without a thematic/taxonomic structure. Overall, our results provide
evidence that cross-cultural differences in similarity judgment are
related to linguistic statistics patterns (which also vary across
cultures).

There are some important limitations of our approach. While we discuss
cross-cultural variability at the level of countries or larger world
areas, these are not cultural monoliths. For convenience, we
operationalize culture at the level of country, based on where
participants were raised. It is an open question whether performance in
our participant populations (of relatively young and well-educated
adults) is representative of the broader country. This is especially
true for societies with substantial ethnic and cultural variation such
as the US. We expect that our data is likely to underestimate variation
both within and between the countries we sample from.

>>>>

Data collection and analyses for this study were pre-registered, and the
pre-registration is available at {[}blinded for review{]}. We recruited
200 participants from the US, 200 from mainland China, and 199 from
Vietnam. All participants were recruited through snowball sampling,
which, in the US was seeded with student email lists at a large
university, in China (CN) with student groups on WeChat, and in Vietnam
(VN) with Vietnam-based student groups on Facebook. US participants were
compensated with \$5 gift certificates (USD), CN participants received
25CNY through WeChat credit transfer, and VN participants received
50,000₫ (VND) in phone credit,

@@

All stimuli and scripts for the experiment and analysis are available at
{[}blinded for review{]}. We collected and adapted stimuli from previous
studies to create test triads consisting of a cue with one thematic and
one taxonomic match option. For example, ``cow,'' ``grass,'' and
``chicken,'' where ``cow'' is the cue, ``grass'' is the thematic match,
and ``chicken'' the taxonomic match. We included 105 such triads, a
superset including triads pulled from supplemental information and
in-text examples across the literature, and others that we adapted or
created . We selected triads on the basis of cultural familiarity within
the US, Vietnam, and China. The triads were translated from English to
Vietnamese and Mandarin by fluent bilingual speakers of each language,
and back-translated to English by another fluent bilingual who was naive
to the original English version. The back-translations were checked
against the original English to identify ambiguities in the translation,
which were resolved through discussion and selection of alternative
terms when relevant.

@@

\begin{CJK*}{UTF8}{gbsn}
Each participant completed all 105 triads in blocks of 21 trials at a time (10 test triads, 10 filler triads, and 1 attention check per page), by selecting the match most related to the cue. We elicited similarity match responses by asking "Which thing is most closely related to the bolded [cue] item?" In Mandarin Chinese, we used the phrasing for "most closely related" reported by Ji et al. (2004): "关联". And in Vietnamese, we translated this as "liên quan nhất".
The test triads were presented with 105 filler triads mixed in, to obscure the taxonomic-thematic two-answer forced choice structure of the test stimuli and reduce the likelihood that participants would become aware of the design.
The filler triads were groups of three semantically related words, but where the match options were not distinguished by thematic vs. taxonomic similarity, for example, the cue "bird" with match options "lizard" and "toad." Additionally, we included 11 attention check trials, which were formatted similarly to the test and filler triads but included an instruction instead of a cue item, e.g., "Choose wife" with match options "wife" and "husband." In total, each participant completed 210 similarity judgments (105 test triads and 105 fillers) and 11 attention check questions, with all items presented in randomized orders that varied between subjects.
\end{CJK*}

@@

In sum, we replicated the differences documented by Ji et al. (2004)
between the US and China. However, we did not find that Vietnamese
participants grouped particularly with Chinese participants. Instead
they were intermediate between the two. Accordingly, we did not find
strong support for the hypothesis that there are overall biases toward
thematic responding across Asian cultural contexts broadly.

@@

First, to test whether variation in language statistics can explain
differences in similarity judgments, we use a mixed-effects logistic
regression fit individually for each country. Our models predicted
responses by individual participants to particular triad (0=taxonomic or
1=thematic) with fastText predictions (proportion of cosine distance) as
a fixed effect and participant and triad as random effects. If language
statistics contribute to the differences in similarity judgments, we
would expect each language corpus to be a good predictor for similarity
judgments in the corresponding population.

@@

If language statistics are able to predict meaningful culture-specific
variation in similarity judgments (rather than just consistency across
cultures), we would expect each corpus to be the best predictor of its
corresponding culture compared to the other two corpora. We directly
compared the corpus models by including all three corpus predictors as
fixed effects in three mixed-effect regressions (predicting US, VN, and
CN responding) with the same random effects as above.

For US responding: only the English (EN) corpus was a significant
predictor{[}\^{}2{]}. EN corpus: \(\beta\) = -6.96, \(\chi^2\)(1) =
16.57, p \textless{} .001. VI corpus: \(\beta\) = -2.22, \(\chi^2\)(1) =
3.73, p = 0.054. ZH corpus: \(\beta\) = -3.18, \(\chi^2\)(1) = 3.37, p =
0.066.

For Vietnamese participants' responses, only the Vietnamese (VI) and
Mandarin (ZH) corpora were significant predictors\footnote{With the
  preregistered exclusion criteria, all three corpora were significant
  predictors.}. EN corpus: \(\beta\) = -2.98, \(\chi^2\)(1) = 2.58, p =
0.108. VI corpus: \(\beta\) = -2.75, \(\chi^2\)(1) = 4.84, p = 0.028. ZH
corpus: \(\beta\) = -4.39, \(\chi^2\)(1) = 5.49, p = 0.019.

For responses by Chinese participants, only the Mandarin (ZH) and
English (EN) corpus were significant predictors. EN corpus: \(\beta\) =
-3.32, \(\chi^2\)(1) = 7.18, p = 0.007. VI corpus: \(\beta\) = -1.59,
\(\chi^2\)(1) = 3.63, p = 0.057. ZH corpus: \(\beta\) = -5.31,
\(\chi^2\)(1) = 17.69, p \textless{} .001.

We observed some language specificity in this analysis (Figure 2). The
English corpus was the best predictor for US responding, and the
Mandarin corpus was the best predictor for Chinese responding. While
this is not the case with the Vietnamese corpus and Vietnamese
participants' responding, the Vietnamese corpus was still a significant
predictor for this responding. These results support our hypothesis that
specific language statistics can predict cross-cultural variation in
similarity judgment.

@@

However, in all cultural contexts, adding the other two corpora produced
a significantly better fit than the identical model with only the
corresponding corpus included as a predictor (US responses:
\(\chi^2\)(2) = 7.93, p = 0.019; VN responses: \(\chi^2\)(2) = 13.72, p
= 0.001; CN responses: \(\chi^2\)(2) = 10.56, p = 0.005). This analysis
suggests that culture-specific input to similarity judgments (as proxied
by language statistics) do not fully explain cross-cultural differences
in similarity judgment, whether due to noise or sparsity in the
estimates of language statistics or true additional variance picked up
by embeddings from other corpora.

@@

One possible concern with our approach is that our models are only able
to account for variation in similarity judgments because the word
embedding models pick up on cross-cultural differences in the
conceptualization of similarity. Perhaps some aspect of the geometry of
the embedding space specifically reflects these differences rather than
capturing more item-specific patterns of co-occurrence.

If this is the case, the embedding models should be less accurate at
predicting responding for filler items (such as whether ``tomorrow'' or
``yesterday'' is more related to ``today'') because these items do not
have a thematic/taxonomic structure. Additionally, even when one option
might be more related to the cue, that relationship is not systematic
throughout the set of fillers. Alternatively, if the embedding models
capture similarity judgments more generally (beyond thematic/taxonomic
preferences), they should predict judgments for the unstructured filler
items as well as they do for our experimental items.

For each filler triad, we randomly assigned one of the responding
options as `Word1'. Running a mixed-effects logistic regression to
predict responding (Word1 or Word2) with country as a fixed effect and a
random effect structure as above in Question 1, we found no effect of
country on filler responding (\(\chi^2\)(2) = 0.74, p = 0.692).

Using the same mixed-effects logistic regression structure as the single
corpus models in Question 2, we predicted responses (1=word1 or 0=word2)
in each cultural context with the corresponding corpus predictor as a
fixed effect, and participant and triad as random effects. In all
cultural contexts, the corresponding corpus was a significant predictor
of responding, with p \textless{} 0.05 and \(\beta\) from -9.59 to
-3.14. These results show that the word embedding models predict general
similarity in addition to structured thematic/taxonomic similarity.

To investigate whether language statistics predict taxonomic/thematic
triads and filler items differently, we compared the variance accounted
for by the models of the two different types of items. We ran two
mixed-effects logistic regression models that predict responding across
cultural contexts (1=thematic or word1, 0=taxonomic or word2), with the
corresponding corpus predictor as a fixed effect and participant and
item as random effects. Consistent with our results above, the
corresponding corpus was a significant predictor for responding to both
triads and filler items (triads: \(\beta\) = -1.79, \(\chi^2\)(1) =
75.58, p \textless{} .001, fillers: \(\beta\) = -2, \(\chi^2\)(1) =
70.89, p \textless{} .001). Importantly, we found comparable conditional
\(R^2\) values when the corresponding corpus was used to predict only
triad items (\(R^2\) = 0.3) and only filler items (\(R^2\) = 0.26). This
result provides evidence for our view that word embeddings index general
tendencies in similarity judgment, beyond a specific dimension of
thematic/taxonomic relations.

@@

Do members of different cultures vary in their conceptualization of
similarity? Cross-cultural differences in similarity judgment would seem
to suggest this conclusion. In this paper, we consider whether
statistics of the environment (as indexed by language statistics) can
account for cross-cultural differences in a classic similarity judgment
paradigm. We replicated the previously documented contrast between
English speakers in the US and Mandarin Chinese speakers from East Asia
(mainland China, Taiwan, Hong Kong, and Singapore): Chinese participants
preferred thematic relations to a greater extent compared to US
participants. Our sample of Vietnamese participants showed an
intermediate preference but were not significantly different from either
Chinese or US participants. This finding suggests some limitations on
the generality of the cultural account, which proposes that
thematic/taxonomic similarity preferences align with East Asian and
Western tendencies toward holistic and analytic processing,
respectively.

In contrast, we found some support for the environmental statistics
account: word embeddings based on language-specific corpora were a good
predictor for the corresponding country's similarity judgments, even
when other corpus statistics were included, and even for filler triads
without a thematic/taxonomic structure. Overall, our results provide
evidence that cross-cultural differences in similarity judgments are
related to patterns in linguistic statistics (which also vary across
cultures).

There are some important limitations of our approach. While we discuss
cross-cultural variability at the level of countries or larger world
areas, these are not cultural monoliths. For convenience, we
operationalize culture at the level of country, based on where
participants were raised and had lived for the large majority of their
lives. It is an open question whether performance in our participant
populations (of relatively young and well-educated adults) is
representative of the broader country. This is especially true for
societies with substantial ethnic and cultural variation such as the US.
We expect that our data is likely to underestimate variation both within
and between the countries we sample from.
